### 2025-02-24 paper/sections/Abstract.tex
initial version of paper without appendix

\begin{abstract}
    salam.
    khobi.
\end{abstract}

>>>>

\begin{abstract}
    This paper explores multilingual emotion classification across binary classification, intensity estimation, and cross-lingual detection tasks. To address linguistic variability and limited annotated data, we evaluate various deep learning approaches, including transformer-based embeddings and traditional classifiers. After extensive experimentation, language-specific embedding models were selected as the final approach, given their superior ability to capture linguistic and cultural nuances. Experiments on high- and low-resource languages demonstrate that this method significantly improves performance, achieving competitive macro-average F1 scores. Despite these advances, challenges remain due to limited annotated data in underrepresented languages and the complexity of nuanced emotional expressions. The study highlights the need for robust, language-aware emotion recognition systems and emphasizes future directions, including expanding multilingual datasets, refining models.
\end{abstract}

### 2025-02-24 paper/sections/Conclusion.tex
initial version of paper without appendix

\section{Conclusion}
This study investigated multilingual emotion classification, addressing the challenges of detecting and analyzing emotions across diverse linguistic contexts. By structuring the problem into three tracks: binary emotion classification, intensity estimation, and cross-lingual emotion detection—we evaluated multiple machine learning and deep learning approaches to improve emotion recognition across various languages. Our findings demonstrate the effectiveness of leveraging innovative preprocessing techniques, domain adaptation strategies, and transfer learning in enhancing model performance.

The significance of this research lies in its contribution to multilingual emotion recognition, particularly for low-resource languages. Our methodology, integrating deep learning models such as LSTMs, transformer-based embeddings, and classification techniques like SVM and XGBoost, achieved competitive results in emotion classification tasks. The study underscores the importance of innovative preprocessing and domain adaptation techniques in addressing linguistic diversity.

Despite the promising results, certain limitations persist. The scarcity of annotated data for many underrepresented languages remains a challenge, potentially impacting model generalization. Additionally, while the study explores multiple model architectures, further optimizations could enhance performance, particularly in cross-lingual settings. Future research should focus on expanding high-quality multilingual emotion datasets, exploring zero-shot learning approaches, and improving model interpretability to enhance emotion classification across diverse languages and cultural contexts.

By advancing multilingual emotion recognition, this research contributes to computational linguistics, social media analysis, and human-computer interaction, fostering better understanding of emotional expressions in multilingual communication.

>>>>

\section{Conclusion}
This study presented a comprehensive examination of multilingual multi-label emotion detection, addressing binary classification, intensity estimation, and cross-lingual detection tasks. Our findings indicate that language-specific embedding models, when paired with classifiers such as SVM and XGBoost, offer a robust approach to capturing the nuanced linguistic and cultural features inherent in diverse textual data. The experimental results, measured in competitive macro-average F1 scores, underscore the potential of these tailored models to bridge performance gaps, particularly in low-resource languages where data scarcity and complex grammatical structures present significant challenges.

The significance of this research lies in its demonstration that integrating innovative preprocessing techniques with state-of-the-art embedding models can lead to substantial improvements in emotion recognition performance. This has broad implications for applications in social media analysis, behavioral research, and other domains where understanding nuanced emotional expressions is crucial.

Nonetheless, certain limitations persist. The scarcity of annotated data in underrepresented languages and the inherent complexity of nuanced emotional expressions continue to impede model performance. Future research should focus on expanding multilingual datasets, refining preprocessing methods, and exploring novel architectures to enhance model generalization and adaptability. Additionally, fine-tuning multilingual models specifically for low-resource languages could significantly improve emotion detection accuracy in these contexts. Such efforts will be instrumental in advancing the field of multilingual emotion analysis and in developing more robust, language-aware emotion recognition systems.

% This study investigated multilingual emotion classification, addressing the challenges of detecting and analyzing emotions across diverse linguistic contexts. By structuring the problem into three tracks: binary emotion classification, intensity estimation, and cross-lingual emotion detection—we evaluated multiple machine learning and deep learning approaches to improve emotion recognition across various languages. Our findings demonstrate the effectiveness of leveraging innovative preprocessing techniques, domain adaptation strategies, and transfer learning in enhancing model performance.

% The significance of this research lies in its contribution to multilingual emotion recognition, particularly for low-resource languages. Our methodology, integrating deep learning models such as LSTMs, transformer-based embeddings, and classification techniques like SVM and XGBoost, achieved competitive results in emotion classification tasks. The study underscores the importance of innovative preprocessing and domain adaptation techniques in addressing linguistic diversity.

% Despite the promising results, certain limitations persist. The scarcity of annotated data for many underrepresented languages remains a challenge, potentially impacting model generalization. Additionally, while the study explores multiple model architectures, further optimizations could enhance performance, particularly in cross-lingual settings. Future research should focus on expanding high-quality multilingual emotion datasets, exploring zero-shot learning approaches, and improving model interpretability to enhance emotion classification across diverse languages and cultural contexts.

% By advancing multilingual emotion recognition, this research contributes to computational linguistics, social media analysis, and human-computer interaction, fostering better understanding of emotional expressions in multilingual communication.

### 2025-02-24 paper/sections/ExperimentalSetup.tex
initial version of paper without appendix

Preprocessing of the dataset was performed using the \texttt{clean-text} library. The preprocessing pipeline included standardizing the text by lowercasing all characters, removing special characters and punctuation, and normalizing tokens through contraction expansion and the elimination of redundant whitespace. These steps ensured the data was clean and consistent across all subsets. Details of the preprocessing pipeline are provided in the supplementary material (Appendix A). Preprocessing was applied consistently to the training, validation, and test datasets to avoid introducing biases or inconsistencies. Such preprocessing steps have been shown to improve the performance of natural language processing (NLP) models by reducing noise and simplifying the input representations \citep{Zhang2020DataPrep}.

@@

Hyperparameter tuning was conducted using \texttt{Optuna}, an advanced framework for automated hyperparameter optimization \citep{Akiba2019Optuna}. The hyperparameters tuned included the learning rate, batch size, dropout rate, and the number of transformer layers to fine-tune. A Bayesian optimization strategy was employed, balancing exploration of the hyperparameter space and exploitation of promising configurations. The performance of each hyperparameter configuration was evaluated on the validation set, and the best configuration was selected based on the performance metric of interest.

@@

The model training process relied on Binary Cross-Entropy (BCE) as the loss function for Track A and C, and Cross-Entropy for Track B due to its suitability for classification tasks. Given the unbalanced dataset, a weighted loss approach was employed to ensure that the model adequately learned from all classes. Optimization was performed using the \texttt{AdamW} optimizer, which improves upon the standard Adam optimizer by decoupling weight decay and learning rate updates \citep{Loshchilov2019AdamW}. To further enhance training stability and convergence, a cosine annealing learning rate scheduler with restarts was employed. This scheduling approach helped adaptively reduce the learning rate over time, facilitating better exploration of the loss landscape and improving generalization. The model was trained for a fixed number of epochs, and early stopping was used to terminate training if the validation performance plateaued, thus avoiding overfitting.

@@

The implementation of the experiments utilized several state-of-the-art tools and libraries. The deep learning models were implemented and trained using \texttt{PyTorch}, which provides a flexible and high-performance framework for machine learning research \citep{Paszke2019PyTorch}. For data manipulation and evaluation metrics, \texttt{Scikit-Learn} was employed \citep{Pedregosa2011ScikitLearn}. Gradient boosting models were benchmarked using \texttt{XGBoost}, which is widely regarded for its efficiency and scalability \citep{chen2016xgboost}. Pre-trained transformer models were fine-tuned using \texttt{Hugging Face Transformers}, a library that offers robust implementations of transformer-based architectures \citep{Wolf2019HuggingFace}. These tools and libraries are well-regarded in the machine learning community and were chosen for their reliability and performance.

@@

All experiments were run on a Kaggle Tesla P100 GPU, which provided the necessary computational power to efficiently train and evaluate the models. The use of a Tesla P100 allowed for accelerated processing, making it feasible to leverage complex models and extensive hyperparameter tuning within a reasonable timeframe. This setup ensures that the experiments are reproducible and can be replicated by other researchers with access to similar hardware.

\subsection{Postprocessing}

Postprocessing involved thresholding the model's output probabilities to convert them into binary class predictions. A default threshold of 0.5 was used, but sensitivity analyses were conducted to evaluate the impact of varying the threshold. This ensured that the model's performance was robust across different operating points and provided interpretable results for downstream tasks. Minimal additional postprocessing was applied to maintain the integrity of the model's predictions.

>>>>

Preprocessing of the dataset was performed using the \texttt{clean-text} library. The preprocessing pipeline involved multiple steps to clean and standardize the text data. Initially, all text was converted to lowercase, and unnecessary whitespace was removed to eliminate redundancy. Special characters, URLs, and emojis were filtered out using regular expressions, with emojis being replaced by their corresponding textual descriptions (e.g., \smiley $\rightarrow$ ``smiling face''). Punctuation was also removed, and  tokenization was performed using language-specific tokenizers to ensure optimal segmentation, and stopwords were removed to further reduce noise. These steps ensured the data was clean and consistent across all subsets. Preprocessing was applied consistently to the training, validation, and test datasets to avoid introducing biases or inconsistencies. Such preprocessing steps have been shown to improve the performance of NLP models by reducing noise and simplifying the input representations \citep{Zhang2020DataPrep}.

@@

Hyperparameter tuning was conducted using \texttt{Optuna}, an advanced framework for automated hyperparameter optimization \citep{Akiba2019Optuna}. The hyperparameters tuned included the hyperparameters of SVM and XGBoost. A Bayesian optimization strategy was employed, balancing exploration of the hyperparameter space and exploitation of promising configurations. The performance of each hyperparameter configuration was evaluated on the validation set, and the best configuration was selected based on the performance metric of interest.

@@

The model training process relied on Binary Cross-Entropy (BCE) as the loss function for Track A and C, and Cross-Entropy for Track B due to its suitability for classification tasks. Given the unbalanced dataset, a weighted loss approach was employed to ensure that the model adequately learned from all classes. Optimization for fine-tuning deep learning models was performed using the \texttt{AdamW} optimizer, which improves upon the standard Adam optimizer by decoupling weight decay and learning rate updates \citep{Loshchilov2019AdamW}. To further enhance training stability and convergence, a cosine annealing learning rate scheduler with restarts\cite{loshchilov2017sgdrstochasticgradientdescent} was employed. This scheduling approach helped adaptively reduce the learning rate over time, facilitating better exploration of the loss landscape and improving generalization. The model was trained for a fixed number of epochs, and early stopping was used to terminate training if the validation performance plateaued, thus avoiding overfitting.

@@

The implementation of the experiments utilized several state-of-the-art tools and libraries. The deep learning models were implemented and trained using \texttt{PyTorch}, which provides a flexible and high-performance framework for machine learning research \citep{Paszke2019PyTorch}. For data manipulation and evaluation metrics, \texttt{Scikit-Learn} was employed \citep{Pedregosa2011ScikitLearn}. Gradient boosting models were benchmarked using \texttt{XGBoost}, which is widely regarded for its efficiency and scalability \citep{chen2016xgboost}. Pre-trained transformer models were fine-tuned using \texttt{Hugging Face Transformers}, a library that offers robust implementations of transformer-based architectures \citep{Wolf2019HuggingFace}. Additionally, the \texttt{Sentence Transformers}\cite{reimers-2019-sentence-bert}\cite{reimers-2020-multilingual-sentence-bert} library was used to load embedding models. These tools and libraries are well-regarded in the machine learning community and were chosen for their reliability and performance.

@@

All experiments were conducted using a Kaggle Tesla P100 GPU, which delivered the computational power required to efficiently train and evaluate the models. The Tesla P100 enabled faster processing, allowing us to effectively utilize complex models and perform extensive hyperparameter tuning within a practical timeframe. This setup ensures that the experiments are reproducible and can be replicated by other researchers with access to comparable hardware.

% \subsection{Postprocessing}

% Postprocessing involved thresholding the model's output probabilities to convert them into binary class predictions. A default threshold of 0.5 was used, but sensitivity analyses were conducted to evaluate the impact of varying the threshold. This ensured that the model's performance was robust across different operating points and provided interpretable results for downstream tasks. Minimal additional postprocessing was applied to maintain the integrity of the model's predictions.

### 2025-02-24 paper/sections/RelatedWork.tex
initial version of paper without appendix

% \section{Related Works}
Several studies have addressed multi-label emotion detection, focusing on various model architectures and preprocessing strategies. In \cite{wang2016dimensional}, a comparative analysis of LSTM and CNN models was performed, demonstrating that hybrid architectures could enhance the F1-score performance of sentiment classification tasks. This research serves as a foundation for designing a model suited to the task at hand.

>>>>

% \section{Related Works}
% Several studies have addressed multi-label emotion detection, focusing on various model architectures and preprocessing strategies. In \cite{wang2016dimensional}, a comparative analysis of LSTM and CNN models was performed, demonstrating that hybrid architectures could enhance the F1-score performance of sentiment classification tasks. This research serves as a foundation for designing a model suited to the task at hand.

### 2025-02-24 paper/sections/SystemOverview.tex
initial version of paper without appendix

The preprocessing pipeline involves multiple steps to clean and standardize the text data. Initially, all text is converted to lowercase, and unnecessary whitespace is removed. Special characters, URLs, and emojis are filtered out using regular expressions. Emojis, in particular, are replaced with their corresponding textual descriptions (e.g., \smiley $\rightarrow$ ``smiling face''). Tokenization is performed using language-specific tokenizers to ensure optimal segmentation. Finally, stopwords are removed, and text normalization techniques, such as stemming or lemmatization, are applied where appropriate.

@@

To extract features, we employed a diverse range of models, including Long Short-Term Memory (LSTM) networks, multilingual language models, and Large Language Models (LLMs). The LLMs were fine-tuned using Low-Rank Adaptation (LoRA) \citep{hu2021lora}, a parameter-efficient tuning method that facilitates task-specific adaptation while maintaining computational efficiency.

@@

Following feature extraction, we applied multiple classification algorithms to perform the multi-label classification task. One of the classifiers used was the Multi-Layer Perceptron (MLP), a feedforward artificial neural network capable of modeling complex relationships between the extracted features and the target labels. Additionally, the system employed XGBoost, a gradient boosting framework renowned for its effectiveness in structured data classification \citep{chen2016xgboost}. Furthermore, SVMs were utilized as a classification method due to their ability to operate effectively in high-dimensional feature spaces by identifying optimal hyperplanes for classification \citep{cortes1995support}.

For languages with sufficient pretrained models available using MTEB\footnote{Massive Text Embedding Benchmark}\citep{muennighoff2022mteb}, we identified the best-performing embedding model and paired it with SVM as the classifier. This approach leverages the strengths of the pretrained embedding models in capturing language-specific nuances, while the SVM classifier ensures robust performance for multi-label classification. On the other hand, for languages with limited pretrained resources, we utilized the multilingual embedding model "Multilingual E5 large instruct" \citep{wang2024multilingual} in combination with XGBoost as the classifier. The model, designed to generalize across diverse languages, enabled the system to maintain high performance even in resource-constrained settings.

>>>>

The preprocessing pipeline involves several steps to clean and standardize the text data. These include converting text to lowercase, removing unnecessary whitespace, filtering out special characters, URLs, and emojis by replacing with their textual description, normalizing tokens, performing language-specific tokenization, and removing stopwords. These steps ensure the data is consistent and suitable for NLP\footnote{natural language processing} tasks.

@@

To extract features, we employed a diverse range of models, including LSTM\footnote{Long Short-Term Memory} networks, MLMs\footnote{Multilingual Language Models}, and LLMs\footnote{Large Language Models}. The LLMs were fine-tuned using LoRA\footnote{Low-Rank Adaptation} \citep{hu2021lora}, a parameter-efficient tuning method that facilitates task-specific adaptation while maintaining computational efficiency.

@@

Following feature extraction, we applied multiple classification algorithms to perform the multi-label classification task. One of the classifiers used was the MLP\footnote{Multi-Layer Perceptron}, a feedforward artificial neural network capable of modeling complex relationships between the extracted features and the target labels. Additionally, the system employed XGBoost, a gradient boosting framework renowned for its effectiveness in structured data classification \citep{chen2016xgboost}. Furthermore, SVMs were utilized as a classification method due to their ability to operate effectively in high-dimensional feature spaces by identifying optimal hyperplanes for classification \citep{cortes1995support}.

For languages with sufficient pretrained models available using MTEB\footnote{Massive Text Embedding Benchmark}\citep{muennighoff2022mteb}, we identified the best-performing embedding model and paired it mainly with SVM as the classifier. This approach leverages the strengths of the pretrained embedding models in capturing language-specific nuances, while the SVM classifier ensures robust performance for multi-label classification. On the other hand, for languages with limited pretrained resources, we utilized the multilingual embedding model ``Multilingual E5 large instruct'' \citep{wang2024multilingual} in combination with XGBoost as the classifier. The model, designed to generalize across diverse languages, enabled the system to maintain high performance even in resource-constrained settings.

### 2025-02-24 paper/sections/introduction.tex
initial version of paper without appendix

\section{Introduction}
The analysis and processing of emotions from textual data have become crucial in understanding human communication across different languages and cultures. This study focuses on the detection and classification of emotions across diverse linguistic contexts, spanning regions from South America to East Asia. Our objective is to categorize emotions into key dimensions, namely sadness, anger, fear, joy, and surprise, while considering cross-lingual variations and linguistic complexities.

>>>>

\section{Introduction}
The analysis and processing of emotions from textual data have become crucial in understanding human communication across different languages and cultures. This study focuses on the detection and classification of emotions across diverse linguistic contexts, spanning regions from South America to East Asia. Our objective is to categorize emotions into key dimensions, namely sadness, anger, fear, disgust, joy, and surprise, while considering cross-lingual variations and linguistic complexities.

### 2025-02-24 paper/tables/MainResults.tex
initial version of paper without appendix

% Multilingual-E5-large-instruct \citep{wang2024multilingual}
% AraBert Large Algerian \citep{arabert}

@@

            \textbf{Language}                       & \textbf{Our Model Name}                                                                      & \textbf{Ours}                & \textbf{Best}                & \textbf{Rank}               & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} \\

@@

            Chinese(chn)                            & ZPoint Large\footnote{\url{https://huggingface.co/iampanda/zpoint_large_embedding_zh}} + SVM & 56.65                        & 70.94                        & 30/41                       & 48.47         & 72.24         & 51.86             & 19/27         & 56.65         & 56.65         & 55.23             & 7/15          \\

@@

            English(eng)                            & \citep{zhang2025jasperstelladistillationsota}                                                & 73.97                        & 82.30                        & 35/96                       & 68.81         & 84.04         & 64.15             & 28/43         & 73.97         & 82.07         & 65.58             & 4/17          \\

@@

            Pt\footnote{Portuguese} Brazilian(ptbr) & \citep{souza2020bertimbau} + SVM                                                             & 47.99                        & 68.33                        & 27/42                       & 38.20         & 71.00         & 46.72             & 22/25         & 47.99         & 68.36         & 51.60             & 8/14          \\
            Pt Mozambican(ptmz)                     & \citep{wang2024multilingual}   + SVM                                                         & 50.08                        & 54.77                        & 6/36                        & —             & —             & —                 & —             & 50.08         & 55.54         & 40.44             & 3/14          \\

@@

@@

>>>>

@@

            \textbf{Language}    & \textbf{Our Model}                                                     & \textbf{Ours}                & \textbf{Best}                & \textbf{Rank}               & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} & \textbf{Ours} & \textbf{Best} & \textbf{Baseline} & \textbf{Rank} \\

@@

            Chinese(chn)         & \citep{zpoint-large-embedding-zh} + SVM                                & 56.65                        & 70.94                        & 30/41                       & 48.47         & 72.24         & 51.86             & 19/27         & 56.65         & 56.65         & 55.23             & 7/15          \\

@@

            English(eng)         & \citep{zhang2025jasperstelladistillationsota}  + SVM                   & 73.97                        & 82.30                        & 35/96                       & 68.81         & 84.04         & 64.15             & 28/43         & 73.97         & 82.07         & 65.58             & 4/17          \\

@@

            Pt* Brazilian(ptbr)  & \citep{souza2020bertimbau} + SVM                                       & 47.99                        & 68.33                        & 27/42                       & 38.20         & 71.00         & 46.72             & 22/25         & 47.99         & 68.36         & 51.60             & 8/14          \\
            Pt* Mozambican(ptmz) & \citep{wang2024multilingual}   + SVM                                   & 50.08                        & 54.77                        & 6/36                        & —             & —             & —                 & —             & 50.08         & 55.54         & 40.44             & 3/14          \\

@@

            % \\
            % \bottomrule

@@

    \\
    \footnotesize{\textit{*: pt=Portuguese}}
